Parsing Postfix log files

### 2007-11-26 doc/logparser-acronyms.tex
Dealt with some XXX items.

\newacronym{IP}{Internet Protocol address}{description={
    The Internet Protocol~\cite{Wikipedia-ip} is the protocol used to
    communicate between computers on the Internet.  An IP address is a
    unique address assigned to a computer on the Internet, allowing it to
    communicate with other computers on the Internet.  For more information
    see~\cite{Wikipedia-ip-address}.
}}

>>>>

\newacronym{IP}{Internet Protocol}{description={
    The Internet Protocol~\cite{Wikipedia-ip} is the protocol used to
    communicate between computers on the Internet.  An IP address is a
    unique address assigned to a computer on the Internet, allowing it to
    communicate with other computers on the Internet.  For more information
    see~\cite{Wikipedia-ip-address}.
}}

### 2007-11-26 doc/logparser.tex
Dealt with some XXX items.

Postfix uses several restriction lists (each containing zero or more
restrictions) to decide the fate of an \SMTP{} command, one list for each
stage of the \SMTP{} conversation: client connection, HELO command, MAIL
FROM command, RCPT TO command (possibly multiple times), DATA command, and
end of data.  By default restriction lists will not be evaluated until the
first RCPT TO command is received, because some clients don't deal with
earlier rejections properly; a side benefit of this delay is that Postfix
has more information available when logging the rejection.\footnote{The
\texttt{smtpd\_delay\_reject} parameter in Postfix's configuration controls
this behaviour.}  Postfix uses simple lookup tables as the deciding factor
for some restrictions, e.g.\ XXX IMPROVE THIS EXPLANATION OR REMOVE IT \\
\tab{}\texttt{check\_client\_access~cidr:/etc/postfix/client\_access}
\footnote{\texttt{cidr}~\cite{cidr_table} is the type of the lookup table.}
\newline checks whether the \IP{} address of the connected client matches
the left hand side of each line in the file and will return the right hand
side of the first matching line.  Other restrictions determine their result
in different ways, e.g.\ \texttt{reject\_rbl\_client rbl.example.com}
checks whether the \IP{} address of the client is present in the listed
\RBL{}\@.

@@

XXX IMPROVE THE NEXT PARAGRAPH BASED ON COMMENTS FROM CARL\@.

XXX REFER TO THE RULES/PARSER SPLIT ELSEWHERE RATHER THAN EXPLAINING IT
AGAIN HERE\@.

XXX``CATEGORISE THE COMPLICATIONS SECTIONS IN WORDS HERE'' --- CARL\@.

@@

Parsing in \LMA{} requires major changes to the code to parse new log lines
or extract additional data.  It does not appear to deal with any of the
complications discussed in sections~\refwithpage{complications}
and~\refwithpage{additional complications}.  The program fails to parse
Postfix 2.2 or 2.3 log files; indeed all it produces is a number of
warning messages from the Perl interpreter, without any other output,
useful or otherwise.  Parsing in \parsername{} is divided in two (see
sections~\refwithpage{rules} and~\refwithpage{parsing-algorithm} for a more
detailed explanation): parsing a new log line is a trivial matter of adding
a new rule, but code changes may be required if new complications
(sections~\refwithpage{complications} and~\refwithpage{additional
complications}) must be dealt with or a new action is required
(section~\refwithpage{actions-in-detail}).

@@

Process scheduling can have a still more confusing effect: quite often the
child mail will be created, delivered and entirely finished with
\textbf{before} the parent logs the creation line!  Thus mails flagged as
coming from an unknown origin cannot be entered into the database when
their final log line is parsed; instead they must be marked as ready for
entry and subsequently entered by the parent mail once it has been
identified.  Quite apart from identifying mail injected in an unknown
fashion, or bugs in the parser, unknown origin mails need to be marked as
such because they lack some data present in the parent mail; they must copy
that data from their parent before being entered in the database.  The
effect of this complication upon the algorithm is discussed more fully in
section~\refwithpage{tracking re-injected mail}.

XXX FIGURE OUT IF I NEED TO COPY DATA ANY MORE, AND DOCUMENT IT IF SO\@.

@@

    \item If the mail is a child mail then it has already been tracked
        (action: commit; flowchart:~20): the data is cleaned up, missing
        data is copied from its parent if necessary, and the child is
        entered in the database (flowchart:~21), then deleted from the
        state tables (flowchart:~22).  The child mail will be removed from
        the parent mail's list of children (flowchart:~23); if this is the
        last child and the parent has also been entered in the database the
        parent will be deleted from the state tables.

>>>>

Postfix uses several restriction lists (each containing zero or more
restrictions) to decide the fate of an \SMTP{} command, one list for each
stage of the \SMTP{} conversation: client connection, HELO command, MAIL
FROM command, RCPT TO command (possibly multiple times), DATA command, and
end of data.  By default restriction lists will not be evaluated until the
first RCPT TO command is received, because some clients don't deal with
earlier rejections properly; a side benefit of this delay is that Postfix
has more information available when logging the rejection.\footnote{The
\texttt{smtpd\_delay\_reject} parameter in Postfix's configuration controls
this behaviour.}  Postfix uses simple lookup tables as the deciding factor
for some restrictions, e.g.\newline
\tab{}\texttt{check\_client\_access~cidr:/etc/postfix/client\_access}
\newline The line above can be broken down as follows:

    \item [check\_client\_access] The name of the restriction to evaluate.

    \item [cidr] The type of the lookup table.

    \item [/etc/postfix/client\_access] The file containing the lookup
        table.

check\_client\_access checks whether the \IP{} address of the connected
client matches the left hand side of each line in the file and returns the
right hand side of the first matching line.  Other restrictions determine
their result by consulting external sources, e.g.\
\texttt{reject\_rbl\_client rbl.example.com} checks whether the \IP{}
address of the client is present in the listed \RBL{}, rejecting the
command if the client is listed.

@@

@@

Parsing in \LMA{} requires major changes to the code to parse new log lines
or extract additional data.  It does not appear to deal with any of the
complications discussed in sections~\refwithpage{complications} (initial
complications) and~\refwithpage{additional complications} (additional
complications).  The program fails to parse Postfix 2.2 or 2.3 log files;
indeed all it produces is a number of warning messages from the Perl
interpreter, without any other output, useful or otherwise.

@@

Process scheduling can have a still more confusing effect: quite often the
child mail will be created, delivered and entirely finished with
\textbf{before} the parent logs the creation line!  Thus mails flagged as
coming from an unknown origin cannot be entered into the database when
their final log line is parsed; instead they must be marked as ready for
entry and subsequently entered by the parent mail once it has been
identified.

XXX MAYBE REMOVE THE NEWPAGE BELOW WHEN I'VE SPLIT THE FLOW CHART\@.

@@

    \item If the mail is a child mail then it has already been tracked
        (action: commit; flowchart:~20): the data is cleaned up, the child
        is entered in the database (flowchart:~21), then deleted from the
        state tables (flowchart:~22).  The child mail will be removed from
        the parent mail's list of children (flowchart:~23); if this is the
        last child and the parent has also been entered in the database the
        parent will be deleted from the state tables.
